---
title: "Data_manipulation_2"
author: "Yiqun Jin"
date: "10/8/2021"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load in the FAS litters Data

```{r}
litters_df = read.csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## select

Choose some columns but not others

```{r}
select(litters_df,group, gd0_weight:gd_of_birth)
```

get rid of certain column

```{r}
select(litters_df, -litter_number)
```

Renaming columns
while identifying the column want to keep

```{r}
select(litters_df, GROUP = group, LITTER_number = litter_number)
```

Keep everything but rename columns

```{r}
rename(litters_df, GROUP = group, LITTER_number = litter_number)
```

Select helpers

```{r}
select(litters_df, starts_with("gd"))
```
move litter_number to the beginning

```{r}
select(litters_df, litter_number, everything())
```

```{r}
relocate(litters_df,litter_number)
```

## filter

remove rows or keep rows that you are interested in retaining

```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd0_weight == 20)
```

```{r}
filter(litters_df, gd_of_birth != 20)
```

```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```
You may also use %in% to detect if values appear in a set

```{r}
filter(litters_df, group  %in% c("Con7","Mod8") )
```

## Mutate
modify some existing variable or create new variable

```{r}
mutate(litters_df, 
       wt_gain = gd18_weight - gd0_weight,
       group = str_to_lower(group))
```

## arrange

```{r}
arrange(litters_df, pups_born_alive)
```

