---
title: "Simple document"
output: html_document
---


```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

## Import some data

I want to import "FAS_litters.csv"

```{r}
litters_df = read_csv("data_import_examples/FAS_litters.csv")

```

I imported the dataset, Now I wnat better name.

```{r}
names(litters_df)
litters_df = janitor::clean_names(litters_df)
```

Now I have better names. Let's look at the dataset.

```{r}
litters_df

head(litters_df)
tail(litters_df)
```


One I used a lot is 

```{r, eval +FALSE}
view(litters_df)
```

Here's "skimr":

```{r}
skimr::skim(litters_df)
```



## Arguments in 'read_csv'
```{r}
litters_df = 
  read_csv(
    "data_import_examples/FAS_litters.csv",
    skip = 5,
    col_names = FALSE,
    na = "Low8"
  )
```


## Parsing columns

```{r}
litters_data = read_csv(file = "./data_import_examples/FAS_litters.csv",
  col_types = cols(
    Group = col_character(),
    `Litter Number` = col_character(),
    `GD0 weight` = col_double(),
    `GD18 weight` = col_double(),
    `GD of Birth` = col_integer(),
    `Pups born alive` = col_integer(),
    `Pups dead @ birth` = col_integer(),
    `Pups survive` = col_integer()
  )
)
tail(litters_data)
```


## Reading from Excel

Reading MLB data

```{r}
mlb11_df = read_excel("data_import_examples/mlb11.xlsx")
```

LotR words is next

```{r}
fellow_df = read_excel("data_import_examples/LotR_Words.xlsx", range = "B3:D6")
```

### Read a SAS file

This is where I tried to read a SAS file

```{r, eval = FALSE}
pulse_df = read_sas("data_import_examples/public_pulse_data.sas7bdat")
```


## why to never use "read.csv"

"read.csv" always import a data frame instead of a table

```{r, eval = FALSE}
litters_df_base = read.csv("data_import_examples/FAS_litters.csv")
litters_df_base
```

## how do I export data

there's a good way!

```{r}
write_csv(fellow_df,"data_import_examples/fellowship_words.csv")
```

P.S. eval = FALSE, do not run the results

