---
title: "data_manipulation"
output: github_document
---

```{r}
library(tidyverse)
library(dplyr) 
library(readxl)
library(haven)
```

## Import some data


```{r}
options(tibble.print_min = 5)

pulse_df = 
  read_sas("./data/public_pulse_data.sas7bdat")
```

```{r}
pulse_df

pulse_tidy = 
  pulse_df %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi"
  )


```

## select

You can specify the columns you want to keep by naming all of them
```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```

You can specify a range of columns to keep

```{r}
select(litters_data, group:gd_of_birth)
```

You can also specify columns you’d like to remove:

```{r}
select(litters_data, -pups_survive)
```

You can rename variables as part of this process:

```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

If all you want to do is rename something, you can use rename instead of select. This will rename the variables you care about, and keep everything else:

```{r}
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

There are some handy helper functions for select; read about all of them using ?select_helpers. I use starts_with(), ends_with(), and contains() often, especially when there variables are named with suffixes or other standard patterns

```{r}
select(litters_data, starts_with("gd"))
```

I also frequently use is everything(), which is handy for reorganizing columns without discarding anything:

```{r}
select(litters_data, litter_number, pups_survive, everything())
```

relocate does a similar thing (and is sort of like rename in that it’s handy but not critical):

```{r}
relocate(litters_data, litter_number, pups_survive)
```

## filter

## mutate

Let's add or change columns

```{r}
mutate(litters_data, weight_change = gd18_weight - gd0_weight)
```

rearrange the data

```{r}
arrange(litters_data, gd_of_birth,gd0_weights)
```

##Pipes

```{r}
litters_data_raw = read.csv("data_import_examples/FAS_litters.csv")
litters_clean_names = janitor::clean_names(litters_data_raw)
litters_selected = select(litters_clean_names, group,pups_survive)
litters_filtered = filter(litters_selected,group == "Con7")

litters_df = 
  read.csv("data_import_examples/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  select(group,pups_survive) %>%
  filter(group == "Con7")
  
```

